Weighing Chances and Use Against Risks and Misuse of Emerging Genome Analysis and Manipulation Technology

### 2022-06-07 draft.tex
Jenny finally provides more input :-)

\subsection{DNA and Identity}
\textit{Identity} is a concept in numerous disciplines, from psychology and pedagogy to sociology, ethnology, social and cultural anthropology, historical and literary studies to philosophy, and beyond that in many trans- and interdisciplinary debates, and has been for decades.
The Oxford English Dictionary defines \textit{identity} as follows: "The sameness of a person or thing at all times or in all circumstances; the condition or fact that a person or thing is itself and not something else; individuality, personality"\cite{oxford}. So, how is DNA related to identity?

Every individual has a unique genetic makeup, their own distinct form of the human genome. 
Furthermore, our DNA remains the same from the first instant of an individual’s existence to his or her last breath\cite{dnaidentity}.
Sociologist Abels describes identity as the answer to the question \textit{who am I} and therefore, identity as a differentiating term\cite{abelsidentity}. Consequently, it is not unjustified to combine both. On the one hand, there is a distinctive combination of DNA sequences that is individual for everyone and consistent for life. 
On the other hand, there is a need for clarification about one's own identity, the will to differentiate oneself from others while remaining consistent within oneself. But what do people hope to gain from their genetic test results?

According to Nordgren, new information can lead to a new self-understanding\cite{nordgren}.
The results of genetic testing allow an insight into irrefutable truths, when executed correctly by qualified personnel. 
In the case of a paternity test, a good example for this would be "I am P’s father". The statement is added as a new aspect of one's own perception and thus as part of one's own identity. Same procedure for the statement "I am an African-American with my roots in Nigeria", based on genetic ancestry tracing.

In a study published in 2018, about 30\% of respondents said they had taken a genetic test to learn how their DNA influences their physical or emotional traits\cite{canceropinion}. 
69 \% of Americans surveyed are very confident that there is a complex genetic code inside our cells that helps determine "who we are"\cite{apgfk}.
Apparently, the link between DNA results and information about one's own characteristics thus seems to have already reached some people. Whether the link was clear from the start or created through targeted marketing, it is probably a combination of both.

@@

\subsection{Commercial Aspect of Identity}

DNA tests are more popular than ever. Due to the decline in the cost of genetic analysis in the last two decades, "direct-to-consumer" genetic screening has become affordable for more and more people\cite{rodney}. 
In 2018, the global direct-to-consumer market was valued at approximately 824.1 million USD and is estimated to grow to 6,364.5 million USD by 2028. 
The rise in the public awareness along with the reduced cost and time required for sequencing owing to the technological advancements is responsible for bolstering the industry demand and aiding in reduction of the global healthcare expenditure\cite{bisresearch}.

\vspace{0.2cm}
\begin{quote}
"Get insights from your DNA, whether it's your ethnicity or personal traits." (Ancestry.com)
\end{quote}
\vspace{0.2cm}
\begin{quote}
"Find out your likelihood of having certain characteristics. See how your DNA affects your hair color, taste preferences and more." (23andme.com)
\end{quote}
\vspace{0.2cm}

Many genetic testing companies appeal in their rhetoric to identity. The previous quotes use words like "characteristics" and "personal traits" to manifest the connection between DNA testing and the customers identity. 
Nordgreen describes this as genetic individualism which is the view that the DNA of an individual gives the individual a unique identity\cite{nordgren}.
This is in accordance with the statements from the previous chapter about DNA test results leading to a new self-understanding.
Furthermore, Nordgren state that modern societies are searching for identity and this rhetoric is a response to that\cite{nordgren}.
In addition to that, Nordgren and Juengst believe that the business models of these companies represent the confluence of three very different currents within contemporary Euro-American culture: the distinctly pre-modern search for a naturalistic understanding of individual identity in a pluralistic world, the thoroughly modern cachet of genomics as a science, and the post-modern emphasis on radical individual self-determination\cite{nordgrenjuengst}. 
The marketing methods of the websites partly explain the growing popularity of DNA tests, but what are the opportunities and risks?

Giving people knowledge of their genetic code, advocates of such services argue, can help them make better decisions about lifestyle, health, and medical care. Also, many families have been reunited due to DNA testing. German television show "Julia Leischik sucht – Bitte melde dich" is an official partner of Ancestry\cite{ancestrywebsite}. In this Real-life documentary series are commissioned by relatives to find people who have not been seen for years, and sets out to find them as soon as she has enough information about the missing person.

The two biggest market players Ancestry and 23andMe have a combined DNA Database Size of 32+ million\cite{ancestrydatabase}\cite{about23andme}. With this amount of data and due to the sensitivity of the content, it should be handled very responsibly. But how does it look in practice?
Laestadius, Rich \& Auer investigated 30 different direct-to-consumer genetic testing websites for two aspects. 
The first aspect being transparency regarding confidentiality and privacy. As a second aspect, the use of data, professional and government bodies created guidelines to promote transparency among these companies was analysed. 
The authors concluded that although most companies met guidelines related to transparency regarding security protocols, storage procedures, and third-party disclosures, only few met guidelines regarding sharing risks from data disclosures. 
Additionally, few companies disclosed how long data would be kept for services or research. 
Use of data for research was frequently mentioned only in privacy policies and terms of service documents, and only two-thirds of companies required an additional consent to use consumer data for health-related research. 
This analysis shows that companies do not consistently meet international transparency guidelines related to confidentiality, privacy, and secondary use of data and are therefore a risk for their customers. 
The authors appeal to clearly inform consumers about specific third parties that will be given access to their data, whether they can have their samples and data deleted at will, and whether their data will be used for anything other than the services they purchased\cite{companydata}.

Another issue is the lack in diversity of the DNA Databases. Chow-White and Duster state that individuals of Asian and African ancestries are underrepresented as well as there are very few DNA samples from Latino and aboriginal peoples. 
If the data is further used in research, as some genome websites allow in their terms and conditions, the production of knowledge about genome variation, medical conditions, and human health is biased. 
This topic will be discussed in more detail in the following chapters.

@@

\subsection{Forensics}
Fingerprints, hair, salvia, and other fluids – the human body leaves traces everywhere. Therefore, its materiality has been a key part of criminal investigations throughout history. 
With the progression of forensic genetics in the last 35 years, many crime offenders have been identified because of the ability to extract DNA profiles\cite{machado}.
 According to Interpol, 63 percent of the asked National Central Bureaus reported having a DNA database (a searchable repository) and they store over 10 million DNA profiles combined\cite{interpol}. 
Germanys Federal Criminal Police Office reports having had over 200,000 hits using DNA analysis in police investigations of crimes from 2000 to 2018\cite{bkastat}.

The controversy around privacy versus surveillance applies to genome analysis as much as to any sensitive technology. 
Germanys BKA describes the evolution of the usage of DNA analysis as going from a supplement to the "classic" identification method of fingerprints to an equal and self-evident means of securing evidence\cite{bkapres}. 
A common claim of advocates for DNA collection is that there is no difference between DNA and fingerprinting in terms of its invasion of an individual's privacy. 
But while fingerprints are a unique identifier, DNA identifies networks of people\cite{cwduster}. DNA data can reveal information about third parties that they have not consented to.
 In the case of the Golden state killer, Criminal DNA databases produced no hits but a match to a distant relative on a genealogy website has led to the identity and therefore the arrest of the perpetrator\cite{goldenstate}. 
In this incident, the misuse of DNA profiles on genealogy websites has led to the prevention of further crimes and therefore has added value for society.

DNA should clearly not be tracked and used carelessly, however statistics from the German Federal Criminal Police Office show that in addition to theft or murder, cases of verbal abuse have also led to hits in DNA analysis\cite{bkastat}. 
What kind of cases make it legitimate to order a genome analysis after an insult is not explained – but the DNA profiles remain in the forensic database. In 2008, the European Court of Human Rights ruled that a retention regime that permits the indefinite retention of DNA records of both convicted and non-convicted (“innocent”) individuals is disproportionate\cite{smarpervsuk}. 
As a consequence, the Protection of Freedoms Act 2012 was implemented to permit the indefinite retention of DNA profiles of most convicted individuals and temporal retention for some first-time convicted minors and innocent individuals on the National DNA Database\cite{protectionoffreedomsact}. 
It is not yet certain whether other countries will follow suit with this data protection directive.

\subsection{Paternity}
After testing for genetic relatedness, non-legal and legal paternity tests are offered most often by direct-to-consumer genetic companies\cite{dnastatista}.
Genetic tests can be helpful in establishing evidence for the parenthood of a person for a case like child custody and support. 
The results of genetic test can also be used as a support for placing a parent’s name on the birth certificate of a child or for settling disputes in child custody laws. However, whether the sampling is done with the consent of all parties tested is not the responsibility of the genetic websites.

@@

\section{Conclusion}
Since the Assyrians and Mendel, a long time has passed. The development of Genome Analysis has become cheaper and is useful in many areas. The police can identify criminals more quickly, doctors can identify illnesses at an early stage and private individuals can learn more about their origins and personality.
Other controversial topics are the general data storage of sensitive DNA data, transparency regarding access by third parties and possibilities for deletion of DNA profiles. In many areas, more emphasis is placed on the collection of data than on the responsible handling of it and the deletion of useless DNA profiles, which makes it potentially easier to misuse the data.

@@

@@

\bibitem{bisresearch} BIS Research,
\textbf{Global Direct-to-Consumer Genetic Testing (DTC-GT) Market} (May, 2019)
https://www.researchandmarkets.com/reports/4771288/global-direct-to-consumer-genetic-testing-dtc. Retrieved May 22th, 2022

\bibitem{about23andme} 23andMe Inc
\textbf{Company Facts}
https://www.23andme.com/about/. Retrieved May, 27th, 2022

\bibitem{companydata} Laestadius, L., Rich, J. \& Auer, P
\textbf{All your data (effectively) belong to us: data practices among direct-to-consumer genetic testing firms} (September 22th, 2016)
https://www.nature.com/articles/gim2016136\#citeas

@@

@@

\bibitem{kpmg} Lauren Friend, Jessica O’Neill, Adrienne Rivlin, Robert Browne,
\textbf{Direct-to-consumer genetic testing} Opportunities and risks in a rapidly evolving market (KPMG 2018) https://assets.kpmg/content/dam/kpmg/xx/pdf/2018/08/direct-to-consumer-genetic-testing.pdf

>>>>

\subsection{DNA and Identity}
\textit{Who am I} is a question of humankind. Many try to find an answer to it - and if not, they will be confronted with opinions of others without being asked. Sociologist Abels describes identity as the answer to the question. 
In addition to that, he proposes the thesis that in advanced modernity we are increasingly asked the question of our own identity. 
As a result, people start to wonder whether they really know or want to know everything about themselves, or whether they have the courage and strength to face uncomfortable truths. 
According to Abels, identity is a term that decides individuals from one another\cite{abelsidentity}.

Identity is a concept in numerous disciplines, from psychology and pedagogy to sociology, ethnology, social and cultural anthropology, historical and literary studies to philosophy, and beyond that in many trans- and interdisciplinary debates, and has been for decades. 
To pursue all approaches would probably require a separate paper.
The Oxford English Dictionary defines identity as follows: "The sameness of a person or thing at all times or in all circumstances; the condition or fact that a person or thing is itself and not something else; individuality, personality"\cite{oxford}. Hence, identity is a constant that applies to an individual's life. 
So, how is DNA related to identity?

As explained in the previous chapter, the human genome consists of about 3.3 billion base pairs. The arragement of sequences in each organism is unique.
Therefore, every individual has a unique genetic makeup, their own distinct form of the human genome. 
DNA remains unchanged throughout all stages of our growth, development, and degeneration. 
The individual’s DNA sequences are stable despite the replacement of chemical elements. 
They persist irrespective of damage to DNA due to random accidents. The sequences do not depend on cognitive abilities or consciousness. 
The Alzheimer’s sufferer who has lost most of her memory has the same genetic base as she had as an infant without self-awareness, or as an adult during the peak of a successful career.
It is the same from the first instant of an individual’s existence to his or her last breath\cite{dnaidentity}.

The connection between identity and the information that DNA contains is close and it is not unjustified to combine both. 
On the one hand, there is a need for clarification about one's own identity, the will to differentiate oneself from others while remaining consistent within oneself throughout life. 
On the other hand, there is a distinctive combination of DNA sequences that is individual for everyone and consistent for life. 
DNA test results could provide new information that anwser the earlier asked question about who you are.
Is this what do people hope to gain from their DNA test results?

According to Nordgren, new information can lead to a new self-understanding\cite{nordgren}.
The results of genetic testing allow an insight into irrefutable truths.
A small example will make this connection clearer.
After a one night stand, a woman becomes pregnant. The only man in question demands a paternity test and, as expected, it turns out that he is the father of the unborn child. 
Now he is sure that the information "I am the father of this child" is true.
The statement is added as a new aspect of his own perception and thus as part of his own identity. 
Another example would be the statement "I am an African-American with my roots in Nigeria", based on genetic ancestry tracing.
DNA testing results allow an insight into the pre-programmed constants of our lives or can extend them by adding irrefutable truths.

@@

\subsection{Commercial Aspect of Identity}
DNA tests are more popular than ever. 
In 2018, the global direct-to-consumer genetic test market was valued at approximately 824.1 million USD and is estimated to grow to 6,364.5 million USD by 2028. This is more than an estimated sevenfold increase over a 10-year period.
There are multiple reasons responsible for bolstering the industry demand.
Firstly, there is the rise in the public awareness. Strategic partnerships with topic-related television programs offer optimal advertising space for suitable target groups.
Best example of this is German television show "Julia Leischik sucht – Bitte melde dich", the official partner of Ancestry\cite{ancestrywebsite}. In this Real-life documentary series are commissioned by relatives to find people who have not been seen for years, and sets out to find them as soon as she has enough information about the missing person. 
This thematically appropriate show provides Ancestry with optimal advertising space.
Secoundly, the technological advancements have lead to reduced cost and time required for sequencing and therfore to affordable prices for private customers\cite{bisresearch}.
The recent decline in the cost of genetic analysis has resulted in a proliferation of companies which provide personal genetic screening directly to consumers and supply them with information about things like genetic ancestry and their relative risk for developing diseases and conditions like colon cancer and Alzheimer’s disease\cite{rodney}.
This chapter focuses on online providers of direct-to-customer DNA testing for the following two reasons: First, marketing activities are easy to view and analyze on the websites. Secound, the terms of use of online providers are listed on the website. The investigation is limited to non-health-related DNA testing as further chapters go into more detail on the healthcare sector.

\vspace{0.2cm}
\begin{quote}
"Get insights from your DNA, whether it's your ethnicity or personal traits." (Ancestry.com)
\end{quote}
\vspace{0.2cm}
In a study published in 2018, about 30\% of respondents said they had taken a genetic test to learn how their DNA influences their physical or emotional traits\cite{canceropinion}. 
69 \% of Americans surveyed are very confident that there is a complex genetic code inside our cells that helps determine "who we are"\cite{apgfk}.
Apparently, the previously introduced connection between identity and the information that DNA contains thus seems to have already reached some people. 
\vspace{0.2cm}
\begin{quote}
"Find out your likelihood of having certain characteristics. See how your DNA affects your hair color, taste preferences and more." (23andme.com)
\end{quote}
\vspace{0.2cm}

As shown in the two quotes, many genetic testing companies appeal in their rhetoric to identity. They use words like "characteristics" and "personal traits" which can be put in connection with one's own identity to manifest the connection between DNA testing and the customers identity. 
The quotes are examples of genetic individualism which is the view that the DNA of an individual gives the individual a unique identity\cite{nordgren}. 
This is in accordance with the statements from the previous chapter about DNA test results leading to a new self-understanding.
Modern societies are searching for identity and this rhetoric is a response to that\cite{nordgren}.
In addition to that, Nordgren and Juengst believe that the business models of these companies represent the confluence of three very different currents within contemporary Euro-American culture: the distinctly pre-modern search for a naturalistic understanding of individual identity in a pluralistic world, the thoroughly modern cachet of genomics as a science, and the post-modern emphasis on radical individual self-determination\cite{nordgrenjuengst}. 
The marketing methods of the websites partly explain the growing popularity of DNA tests, but what are the opportunities and risks?

Prior to the accessibility of DNA testing to the general population, learning more about the history of one's ancestors was a difficult undertaking. 
Of course, this was only possible under the condition that one can show clues such as relatives who are still alive or the names of some ancestors.
However, this does not exclude the possibility of untruths being told about one's own family history or documents being embellished. 
Especially in cases that do not correspond to the image of the social norm, such as illegitimate children or incest, a lot has been covered up.
And at that point, there was no other option but to let the past rest.
However, with direct-to-consumer DNA testing, everyone has the opportunity to learn more about their own family history.
An interesting case which couldn't be solved without DNA testing would be "Chile's stolen children".
During the Pinochet era in Chile, about 20,000 children were adopted by foreign couples of which at least 8,000 children are suspicious cases of illegal child abduction.
The children were stolen from their mothers during the military rule of Gen Augusto Pinochet from 1973 to 1990 and sent abroad for adoption to developed countries, including Holland, the United States, Sweden and Germany.
Some women were forced to sign paperwork they did not understand while others were even told their children had died.
Even after this atrocity is uncovered, the starting point for family reunification would be very tenuous: false information on adoption papers, lack of documentation about the mothers, and the language barrier.
Since 2018, almost 200 mothers have been reunited with their children.
Without DNA testing, these families would not have had a realistic chance of reuniting\cite{chile}. 
Many providers tell of similar success stories of their clients on their websites.

In addition to the opportunity to learn more about one's own family, many geneology test providers also offer paternity testing. After testing for genetic relatedness, non-legal and legal paternity tests are offered most often by direct-to-consumer genetic companies\cite{dnastatista}.
Genetic tests can be helpful in establishing evidence for the parenthood of a person for a case like child custody and support. 
The results of genetic test can also be used as a support for placing a parent’s name on the birth certificate of a child or for settling disputes in child custody laws. However, whether the sampling is done with the consent of all parties tested is not the responsibility of the genetic websites.

Giving people knowledge of their genetic code, advocates of such services argue, can help them make better decisions about lifestyle and health. 
Although there are no representative studies on this, additional information about one's own identity can lead to more self-reflection and thus to a more conscious life. 
Twin and family studies have showed that personality traits are moderately heritable, and can predict various lifetime outcomes, including psychopathology.
Despite considerable efforts over the past several decades and major progress toward unraveling the genetic etiology of personality traits, the genetic variants that influence personality are only beginning to be identified\cite{geneticspersonality}. 
The genetic testing provider 23andMe divides their possible personality traits into three report groups. 
The first report is about "Physical Features" which includes information about Hair Photobleaching, Freckles, Skin Pigmentation, and more. 
The secound report focuses on "Taste and Smell" and reveals more about the customers Cilantro Taste Aversion, Sweet vs. Salty, Asparagus Odor Detection, and. 
The third report is called "Weird and wonderful" and is about Photic Sneeze Reflex, Ability to Match Musical Pitch, Misophonia (hatred of the sound of chewing), and more\cite{23personality}.
Although a large number of personality traits are listed and research into genetic personality research is still in its infancy, there is no reference to the accuracy of the test results on the provider's website. Unfortunately, this is not an isolated case of lack of transparency by a DNA testing provider.

The two biggest market players Ancestry and 23andMe have a combined DNA Database Size of 32+ million\cite{ancestrydatabase}\cite{about23andme}. With this amount of data and due to the sensitivity of the content, it should be handled very responsibly. 
The previous chapter talked in detail about the need for genetic privacy. Do the providers meet the requirement of their customers?
Laestadius, Rich \& Auer investigated 30 different direct-to-consumer genetic testing websites for two aspects. 
The first aspect being transparency regarding confidentiality and privacy. As a second aspect, the use of data, professional and government bodies created guidelines to promote transparency among these companies was analysed. 
The authors concluded that although most companies met guidelines related to transparency regarding security protocols, storage procedures, and third-party disclosures, only few met guidelines regarding sharing risks from data disclosures. 
Additionally, few companies disclosed how long data would be kept for services or research. 
Use of data for research was frequently mentioned only in privacy policies and terms of service documents, and only two-thirds of companies required an additional consent to use consumer data for health-related research. 
This analysis shows that companies do not consistently meet international transparency guidelines related to confidentiality, privacy, and secondary use of data and are therefore a risk for their customers. 
The authors appeal to clearly inform consumers about specific third parties that will be given access to their data, whether they can have their samples and data deleted at will, and whether their data will be used for anything other than the services they purchased\cite{companydata}.

Another issue is the lack in diversity of the DNA Databases. 
Chow-White and Duster state that individuals of Asian and African ancestries are underrepresented as well as there are very few DNA samples from Latino and aboriginal peoples. 
If the data is further used in research, as some genome websites allow in their terms and conditions, the production of knowledge about genome variation, medical conditions, and human health is biased\cite{cwduster}. 
This topic will be discussed in more detail in the following chapters.

@@

\subsection{Forensics}
Fingerprints, blood splatter, hair, skin flake, salvia, and other fluids – the human body leaves traces everywhere. Therefore, its materiality has been a key part of criminal investigations throughout history. 
With the progression of forensic genetics in the last 35 years, many crime offenders have been identified because of the ability to extract DNA profiles\cite{machado}. 
In practice, forensic testing uses DNA sequences to identify an individual for legal purposes. Subsequently, a DNA profile is created for each individualised DNA trace. 
This could be, for example, a biological trace at a crime scene that can be clearly assigned to one person. Other applications would be the identification of a corpse that has become visually unrecognisable due to decay or the assignment of a body part after a natural disaster.
According to Interpol, 63 percent of the asked National Central Bureaus reported having a DNA database (a searchable repository) which store over 10 million DNA profiles combined\cite{interpol}.

The controversy around privacy versus surveillance applies to genome analysis as much as to any sensitive technology. 
Germanys Federal Criminal Police Office describes the evolution of the usage of DNA analysis as going from a supplement to the "classic" identification method of fingerprints to an equal and self-evident means of securing evidence\cite{bkapres}. 
A common claim of advocates for DNA collection is that there is no difference between DNA and fingerprinting in terms of its invasion of an individual's privacy. 
But while fingerprints are a unique identifier, DNA identifies networks of people\cite{cwduster}. DNA discloses information about our family members and ancestors and can therefore reveal information about third parties that they have not consented to.
In the case of the Golden state killer, Criminal DNA databases produced no hits but a match to a distant relative on a genealogy website has led to the identity and therefore the arrest of the perpetrator\cite{goldenstate}. In this incident, the misuse of DNA profiles on genealogy websites has led to the prevention of further crimes and therefore has added value for society.

Bieber, Brenner and Lazer investigated the possibilities of finding criminals through the DNA of their relatives and demonstrate the potential value of kinship analysis for identifying promising leads in forensic investigations.
Indirect genetic relationship analysis using DNA of biological relatives is very useful especially in the previously mentioned cases for mass humanitarian disasters and identification of missing persons. In only a few cases, searches in DNA databases have been conducted to identify close relatives of potential suspects. 
The authors state that the "cold hit" rate - the probability of finding a match between a crime scene sample and someone in the offender database - can be increased by 40\% using kinship analysis.\cite{dnarelatives}

There are a number of factors that help mitigate the intrusions of DNA use, at least in the near future. So far, some specific details can be extracted from DNA: hair color, ethnic origin or whether the suspect is susceptible to certain diseases. 
But it will be several years before authorities can look at the genetic code and put together a composite picture of a person. 
So while one's DNA has the potential to reveal an enormous amount of information about an individual, at this stage, authorities are unable to identify enough genes, their functions, or their expression patterns to actually reveal the information.

DNA should nevertheless be tracked and used carelessly. However, statistics from the German Federal Criminal Police Office show that in addition to theft or murder, cases of verbal abuse have also led to hits in DNA analysis\cite{bkastat}. 
What kind of cases make it legitimate to order a genome analysis after an insult is not explained – but the DNA profiles remain in the forensic database. 
It was not until 2008 that the European Court of Human Rights ruled a retention regime which permits the indefinite retention of DNA records of both convicted and non-convicted (“innocent”) individuals is disproportionate\cite{smarpervsuk}. 
As a consequence, the Protection of Freedoms Act 2012 was implemented in the UK to permit the indefinite retention of DNA profiles of most convicted individuals and temporal retention for some first-time convicted minors and innocent individuals on the National DNA Database\cite{protectionoffreedomsact}. 
It is not yet certain whether other countries will follow suit with this data protection directive.

Unfortunately, some countries have already disregarded the actual purpose of the national DNA databases.
In 2015, Kuwait anncounced the mandate that all 1.3 million citizens and 2.9 million foreign residents have their DNA tested and placed into a database.
The law followed a deadly suicide attack on a mosque in the country's capital\cite{kuweit1}. Fortunately, the country's Constitutional Court ruled that the law violates the constitutional guarantee of personal freedom\cite{kuweit2}.
China is currently criticized by human rights groups and Uighur activists for using its comprehensive DNA database to chase down any Uighurs who resist conforming to the campaign\cite{nytimes}. From these examples it can be seen that national DNA databases can be misused for surveillance, to identify and oppress ethnic groups and they are a useful tool for a potential police state.

Etzioni provides an interesting thought on the boundary between privacy and surveillance. He argues that the benefits for society of storing the DNA of a convicted of or even suspected of breaking the law once in the database outweigh his or her individual privacy rights. 
But if an individual is whether a convict nor a suspect, the balance tips the other way\cite{etzioni}.
As a basic premise, any police agency or state government using DNA profiling and DNA databases should evaluate the political, ethical, and legal implications associated with them, in addition to get their valuable investigative potential.

@@

\section{Conclusion and Outlook}
Since the Assyrians and Mendel, a long time has passed. The development of Genome Analysis has become cheaper and is useful in many areas. 
To give an answer whether the chances overweigh the risks of this technology or if it's the other way around, is not simple. 
The science behind genome analysis should be considered neutral in itself.
Controversial topics are the general data storage of sensitive DNA data, transparency regarding access by third parties and possibilities for deletion of DNA profiles. 
In many areas, more emphasis is placed on the collection of data than on the responsible handling of it and the deletion of useless DNA profiles, which makes it potentially easier to misuse the data.
In order to make the right decision in handling the sensitive data, the valuable investigative potential with its associated political, ethical and legal implications should be assessed in each individual case. A one-size-fits-all approach makes less sense, as many external factors influence the evaluation between societal added value and the individual's privacy. 
In general, every user of DNA tests should be informed about what can be read from the data, what is stored and then decide whether the added value outweighs the potential risk of a leak in his or her individual case.
On condition that the person is given the opportunity to make a choice.

Many genetic test providers attract a larger audience by marketing to give them answers to their questions about their own identity. 
And although, DNA testing for the general public has led to many family reunions, clarification of paternity relationships and insights into one's own origins and personality, there is still not enough transparent information about the transfer of data to third parties. 
Here, the providers are obliged to provide more information about the transfer and to consider the consequences of this transfer for the community.

National DNA databases are useful to solve crimes more quickly, to identify missing persons or corpses or - cooperatively - to stop offences on an international level. In addition to that, there are still methods, like finding criminals through the DNA of their relatives, whose potential has not yet been exhausted.
Caution is advised as soon as a state puts its own political interests above those of its own population.
Reasonably, a monitoring body independent of the state should be introduced to prevent this abuse.

In the medical field, DNA tests can detect and treat diseases at an early stage. Especially hereditary diseases have great potential for this.
However, the enumerated concerns in the areas should be taken into account: Patient Autonomy, Patient Autonomy, Commercial Use, Use of Public Resources, Patient Responsibilities, Psychological Risks. 
In order to prevent genetic discrimination, for example by insurance companies, it is imperative that sensitive data continues to be stored discretely access is tightly controlled as to not enable exploitation of this knowledge.

Eugenics was only touched upon briefly in this paper and is a topic that still needs a lot of discussion.
Although eugenics has been used on a small scale by some civilisations in the past, the current possibilities are not comparable with today's science. 
Perhaps humanity could be brought to another level. However, the implementation of eugenics raises a number of ethical questions that make it impossible to draw clear boundaries and thus to use it in a politically, ethically and legally correct way.

This paper has assessed the opportunities and risks of genome analysis in some relevant areas of application. 
It makes sense to carry out further research in the individual areas and to derive concrete recommendations for action. 
Overall, there is still a lot of need for clarification about the responsibility of data arising from genome analysis.

@@

\bibitem{bisresearch} BIS Research,
\textbf{Global Direct-to-Consumer Genetic Testing (DTC-GT) Market} (May, 2019)
https://www.researchandmarkets.com/reports/4771288/global-direct-to-consumer-genetic-testing-dtc. Retrieved May 22nd, 2022

@@

\bibitem{chile} Jane Chambers
\textbf{Chile's stolen children: 'I was tricked into handing over my baby'} BBC News (September 26th, 2019)
https://www.bbc.com/news/world-latin-america-48929112. Retrieved June 6th, 2022

\bibitem{geneticspersonality} S. Sanchez-Roige, J. C. Gray, J. MacKillop, C.-H. Chen, A. A. Palmer
\textbf{The genetics of human personality} (November 7th, 2017)
https://onlinelibrary.wiley.com/doi/pdf/10.1111/gbb.12439

\bibitem{23personality} 23andMe, Inc.
\textbf{Ancestry +plus Traits Service}
https://www.23andme.com/dna-ancestry/. Retrieved June 6th, 2022

\bibitem{about23andme} 23andMe, Inc.
\textbf{Company Facts}
https://www.23andme.com/about/. Retrieved May, 27th, 2022

\bibitem{companydata} Laestadius, L., Rich, J. \& Auer, P
\textbf{All your data (effectively) belong to us: data practices among direct-to-consumer genetic testing firms} (September 22nd, 2016)
https://www.nature.com/articles/gim2016136\#citeas

@@

\bibitem{dnarelatives} Frederick R. Bieber, Charles H. Brennerand, David Lazer
\textbf{Finding Criminals Through DNA of Their Relatives} (June 2nd, 2006)
https://www.science.org/doi/10.1126/science.1122655

@@

\bibitem{kuweit1} Nick Visser
\textbf{Kuwait To Institute Mandatory DNA Testing For All Residents}, Huffpost (July 14th, 2015)
https://www.huffpost.com/entry/kuwait-mandatory-dna\_n\_7795442. Retrieved June 6th, 2022

\bibitem{kuweit2} Andy Coghlan
\textbf{Kuwait's plans for mandatory DNA database have been cancelled}, NewScientist (October 9th, 2017)
https://www.newscientist.com/article/2149830-kuwaits-plans-for-mandatory-dna-database-have-been-cancelled/. Retrieved June 6th, 2022

\bibitem{nytimes} Sui-Lee Wee
\textbf{China uses DNA to track its people, with the help of American expertise}, The New York Times (Feburary 21st, 2019)
https://www.nytimes.com/2019/02/21/business/china-xinjiang-uighur-dna-thermo-fisher.html

\bibitem{etzioni} David Lazer, ed.
\textbf{DNA and the criminal justice system: The technology of justice} (2004)
